State of the Practice for Lattice Boltzmann Method Software

### 2021-12-29 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
updated LBM paper to include Musubi measurements, up to but not including Section 5.6

The software packages were examined for the presence of artifacts, which were
then categorized by frequency. We have grouped them into common, less common,
and rare artifacts in Table~\ref{artifactspresent}. Common artifacts were found
in 16 to 23 ($>$70\%) of the software packages. Less common artifacts were found
in 8 to 15 (35-70\%) of the software packages. Rare artifacts were found in 1 to
7 ($<$35\%) of the software packages.

@@

All of the top four AHP ranked packages, ESPREsSo, Ludwig, OpenLB, and Palabos,
have each of the commonly found artifacts, except only three of them have a
requirements specification or theory notes. Palabos is the only one of these
four packages that does not have an artifact from this category.

@@

The top four AHP ranked packages have most of the less common artifacts. At the
time of data collection, only one (Palabos) of the four packages did not have a
user manual or guide, but there was a broken link on the package website
indicating that such an artifact might exist. This broken link was later fixed,
but this is not reflected in our data because it was not present at the time of
data collection. Despite the broken link, Palabos does have a detailed and
informative website. Another one (Ludwig) of the top four packages does not
appear to have publicly visible design documentation. A third package (OpenLB)
from the list does not appear to use a version control system. It is possible
that such a system is used by the package since its website notes package
version numbers, but the artifacts do not explicitly state the use of such a
system.

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep{ghezzi1991fundamentals}. The following type of development tools
were explicitly noted in the artifacts or web-pages of the 23 LBM packages that
were assessed. It is likely that other tools, such as debuggers, were used but
are not specified in our sources.

@@

The above tools can verify the correctness of software during its development.
Only two (ESPResSo, Ludwig) of the software packages that were assessed
mentioned using continuous integration tools, like Travis CI. Code editors and
compilers were explicitly noted to have been used by several packages, and were
likely used by all of them. One of the packages (Ludwig) explicitly noted the
use of proprietary unit testing code written in C. Likewise, the use of
proprietary code for verifying the correctness of output was noted by one
(pyLBM) of the developers. It is likely that similar tools were used when
developing other software packages.

@@

The following types of dependencies were explicitly noted in the artifacts or
web-pages of the 23 LBM packages that were assessed. It is possible that other
types of dependencies are part of these software packages, but are not clearly
specified in their artifacts or web sites and because of that they are not
listed here.

@@

Many of the software packages that were assessed were developed by teams of two
or more people. Their work needed to be coordinated and managed. The following
types of project management tools were explicitly noted in the artifacts,
web-pages, or interviews with the developers of the 23 LBM packages that were
assessed. As with development tools and dependencies, it is possible that other
types of project management tools were used to coordinate and manage the
projects but are not specified and because of that they are not listed here.

@@

Collaboration tools are noted as being used when developing the software
projects. Most often email and video conferencing is used. Project management
software was not explicitly mentioned, but it is possible that some of the
projects use such software. Many of the projects are located on GitHub, and its
developers use the platform to help manage their projects, especially bug
related issues. Most of the projects appear to use change tracking and version
control tools. They often use GitHub or GitLab for this. One package
(SunlightLB) uses CVS. Document generation tools are mentioned in the artifacts
of 12 of the projects. The tools Sphinx and Doxygen are explicitly used in this
capacity.

@@

Seven of the software packages contain artifacts outlining the general
development process, like basic instructions on how to contribute. Eleven of the
packages explicitly convey that they would accept outside contributors, but
generally the teams are centralized, often working at the same institution.

@@

Thirteen (57\%) of the 23 software packages use GitHub for managing the project,
including nine (64\%) of the 14 alive packages, and four of the nine (44\%) dead
packages. Two projects (Palabos, WaLBerla) use GitLab. This could be indicative
of a transition to such software development and version control tools for SCS.
Typically there are several simultaneous development branches in these projects.

Documentation was also noted as playing a significant role in the development
process, specifically with on-boarding new developers. A goal of documentation
is to lower the entry barrier for these new contributors. The documentation
provides information on how to get started, orients the user to artifacts and
the source code, and explains how the system works, including the so-called
simulation engine and interface. The use of document generation tools is
mentioned in the artifacts of 12 software packages, and was noted during
interviews with developers. Sphinx and Doxygen are the tools that were
mentioned.

@@

\begin{itemize}
	\item The system must provide descriptive error messages when it encounters unexpected input, including incorrect data types, empty input, and missing files or links. Eighteen of the software packages behaved reasonably when tested with unexpected input.
\end{itemize}

@@

\begin{itemize}
	\item Integrate parallelization tools and techniques to reduce processing time. Consider GPU processing, CUDA, and MPI. Seventeen software packages mentioned parallelization in their artifacts.
	\item The user should be able to choose to process their model on either the CPU or GPU. ESPResSo, lbmpy, lettuce, pyLBM, and TCLB allow for either CPU or GPU processing.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA and HemeLB explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Nineteen of the 23 software packages use such code identifiers.
	\item Use symbolic constants. Twelve of the 23 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code. All of the top ranked packages have meaningful comments in their source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Provide a user manual that identifies all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Summarize the development process that is used. Provide information on how new users can contribute. Identify the development model by name (waterfall, agile, etc.), if appropriate.  Seven of the software packages have some artifacts outlining the general development process. Eleven packages have information on how to contribute. 
	\item Identify the development environment. Eight of the 23 software packages identify the development environment.
	\item Update all relevant documentation when a change to the software is made. This is suggested as good software development practice.
	\item Include notes with all releases. Nine of the packages include release notes.
	\item Communicate all changes within the development team. This suggestion is based on developer comments.
	\item Use continuous integration processes and tools (Bamboo, Jenkins, and Travis CI). This was done by top ranked packages ESPResSo and Ludwig.
	\item Consider peer review processes to assess contributions and ensure the tracking of information. High ranked package ESPResSo uses a peer review process for contributions. 
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx). This was done by all of the top five ranked packages. 
\end{itemize}

>>>>

The software packages were examined for the presence of artifacts, which were
then categorized by frequency. We have grouped them into common, less common,
and rare artifacts in Table~\ref{artifactspresent}. Common artifacts were found
in 16 to 24 ($>$63\%) of the software packages. Less common artifacts were found
in 8 to 15 (30-63\%) of the software packages. Rare artifacts were found in 1 to
7 ($<$30\%) of the software packages.

@@

All of the top four AHP ranked packages, ESPREsSo, Ludwig, Palabos, and OpenLB, have each of the commonly found artifacts, except only three of them have a
requirements specification or theory notes. Palabos is the only one of these
four packages that does not have an artifact from this category.

@@

The top four AHP ranked packages have most of the less common artifacts. At the
time of data collection, only one (Palabos) of the four packages did not have a
user manual or guide, but there was a broken link on the package website
indicating that such an artifact might exist. This broken link was later fixed,
but this is not reflected in our data because it was not present at the time of
data collection. Despite the broken link, Palabos does have a detailed and
informative website. Another one of the top four packages, Ludwig, does not
appear to have publicly visible design documentation. A third package from the list, OpenLB, does not appear to use a version control system. It is possible
that such a system is used by the package since its website notes package
version numbers, but the artifacts do not explicitly state the use of such a
system.

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep{ghezzi1991fundamentals}. The following type of development tools
were explicitly noted in the artifacts or web-pages of the 24 LBM packages that
were assessed. It is likely that other tools, such as debuggers, were used but
are not specified in our sources.

@@

The above tools can verify the correctness of software during its development.
Only three (ESPResSo, Ludwig, Musubi) of the software packages that were assessed mentioned using continuous integration tools, like Travis CI. Code editors and compilers were explicitly noted to have been used by several packages, and were likely used by all of them. One of the packages (Ludwig) explicitly noted the use of proprietary unit testing code written in C. Likewise, the use of proprietary code for verifying the correctness of output was noted by one (pyLBM) of the developers. It is likely that similar tools were used when developing other software packages.

@@

The following types of dependencies were explicitly noted in the artifacts or
web-pages of the 24 LBM packages that were assessed. It is possible that other
types of dependencies are part of these software packages, but are not clearly
specified in their artifacts or web sites and because of that they are not
listed here.

@@

Many of the software packages that were assessed were developed by teams of two
or more people. Their work needed to be coordinated and managed. The following
types of project management tools were explicitly noted in the artifacts,
web-pages, or interviews with the developers of the 24 LBM packages that were
assessed. As with development tools and dependencies, it is possible that other
types of project management tools were used to coordinate and manage the
projects but are not specified and because of that they are not listed here.

@@

Collaboration tools are noted as being used when developing the software
projects. Most often email and video conferencing is used. Project management
software was not explicitly mentioned, but it is possible that some of the
projects use such software. Many of the projects are located on GitHub, and its
developers use the platform to help manage their projects, especially bug
related issues. Most of the projects appear to use change tracking and version
control tools. They often use GitHub or GitLab for this. One package
(SunlightLB) uses CVS. Document generation tools are mentioned in the artifacts
of 13 of the projects. The tools Sphinx and Doxygen are explicitly used in this
capacity.

@@

Seven of the software packages contain artifacts recording the general
development process and status. Eleven of the packages explicitly convey that they would accept outside contributors, but generally the teams are centralized, often working at the same institution.

@@

Thirteen (54\%) of the 24 software packages use GitHub for managing the project,
including nine (60\%) of the 15 alive packages, and four of the nine (44\%) dead
packages. Two projects (Palabos, WaLBerla) use GitLab. This could be indicative
of a transition to such software development and version control tools for SCS.
Typically there are several simultaneous development branches in these projects.

Documentation was also noted as playing a significant role in the development
process, specifically with on-boarding new developers. A goal of documentation
is to lower the entry barrier for these new contributors. The documentation
provides information on how to get started, orients the user to artifacts and
the source code, and explains how the system works, including the so-called
simulation engine and interface. The use of document generation tools is
mentioned in the artifacts of 13 software packages, and was noted during
interviews with developers. Sphinx and Doxygen are the tools that were
mentioned.

@@

\begin{itemize}
	\item The system must provide descriptive error messages when it encounters unexpected input, including incorrect data types, empty input, and missing files or links. Nineteen of the software packages behaved reasonably when tested with unexpected input.
\end{itemize}

@@

\begin{itemize}
	\item Integrate parallelization tools and techniques to reduce processing time. Consider GPU processing, CUDA, and MPI. Eighteen software packages mentioned parallelization in their artifacts.
	\item The user should be able to choose to process their model on either the CPU or GPU. ESPResSo, lbmpy, lettuce, pyLBM, and TCLB allow for either CPU or GPU processing.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA, HemeLB, and Musubi explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Twenty of the 24 software packages use such code identifiers.
	\item Use symbolic constants. Thirteen of the 24 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code. All of the top ranked packages have meaningful comments in their source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Provide a user manual that identifies all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Summarize the development process that is used. Provide information on how new users can contribute. Identify the development model by name (waterfall, agile, etc.), if appropriate.  Seven of the software packages have some artifacts outlining the general development process. Eleven packages have information on how to contribute. 
	\item Identify the development environment. Eight of the 24 software packages identify the development environment.
	\item Update all relevant documentation when a change to the software is made. This is suggested as good software development practice.
	\item Include notes with all releases. Nine of the packages include release notes.
	\item Communicate all changes within the development team. This suggestion is based on developer comments.
	\item Use continuous integration processes and tools (Bamboo, Jenkins, and Travis CI). This was done by top ranked packages ESPResSo and Ludwig.
	\item Consider peer review processes to assess contributions and ensure the tracking of information. High ranked package ESPResSo uses a peer review process for contributions. 
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx). This was done by all of the top five ranked packages. 
\end{itemize}
